% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother


\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother





\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}



 


\KOMAoption{captions}{tableheading}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}


\author{}
\date{}
\begin{document}


Last Updated: Nov 11, 2025

tried new code Download PDF Version

\subsection{Education}\label{education}

PhD Ecology \& Environmental Science \textbar{} {Montana State
University}

2025 - present

\begin{itemize}
\tightlist
\item
  Graduate Committee: Dr.~Robert Peterson (co-chair), Dr.~David Weaver
  (co-chair), Dr.~Sharlene Sing, Dr.~William Wetzel
\end{itemize}

MS Entomology \textbar{} {Montana State University}

2021 - 2024

\begin{itemize}
\tightlist
\item
  Graduate Committee: Dr.~David Weaver (co-chair), Dr.~Tracy Sterling
  (co-chair), Dr.~Robert Peterson
\item
  Graduate Certificate in Applied Statistics
\end{itemize}

BS Biology \textbar{} {University of Minnesota - Twin Cities}

2013 - 2017

\begin{itemize}
\tightlist
\item
  Minor: Marine Biology
\end{itemize}

\noindent\textbf{PhD Ecology \& Environmental Science}
\hfill \textit{Montana State University} \hfill 2025--present
\textbackslash{} Graduate Committee: Dr.~Robert Peterson (co-chair),
Dr.~David Weaver (co-chair), Dr.~Sharlene Sing, Dr.~William Wetzel

\subsection{Publications}\label{publications}

{Check out my~Publications~page}

Strand JR, Peterson RKD, Sterling TM, Weaver DK. 2025. Sensory and
behavioral responses of braconid parasitoids to changes in volatile
emissions induced by wheat stem sawfly (Hymenoptera: Cephidae) larval
feeding in winter wheat and smooth brome. Journal of Insect Science.
25(2). DOI: 10.1093/jisesa/ieaf016

Strand JR, Peterson RKD, Sterling TM, Weaver DK. 2024. Agroecological
importance of smooth brome in managing wheat stem sawfly (Hymenoptera:
Cephidae) via associated braconid parasitoids. Journal of Economic
Entomology. 117(6): 2344--2354. DOI: 10.1093/jee/toae246

\subsection{Honors \& Awards}\label{honors-awards}

{Extreme Biofilms NRT Program SEED Grant} (\$1,500)

2025

\begin{itemize}
\tightlist
\item
  Extreme Biofilms NRT, Montana State Unviersity\\
\item
  Awarded for the proposal ``\emph{Mapping Biocontrol Success: Drone
  NDVI Monitoring of Invasive Toadflax in Post-Fire Montana
  Ecosystems}''
\end{itemize}

Angus and Anne McMillan Family Agriculture Scholarship (\$1,200)

2025

\begin{itemize}
\tightlist
\item
  Montana State University
\end{itemize}

Open Access Author Fund(\$1,736)

2025

\begin{itemize}
\tightlist
\item
  Montana State University Library\\
\item
  Funded the publication of \emph{Strand et al.~2025}
\end{itemize}

NAISMA 2025 Student Scholarship

2025

\begin{itemize}
\tightlist
\item
  National Forum on Biological Control, Annapolis, MD
\end{itemize}

Graduate Land Grant Excellence Award

2024

\begin{itemize}
\tightlist
\item
  The Graduate School, Montana State University
\end{itemize}

Open Access Author Fund(\$3,296)

2023

\begin{itemize}
\tightlist
\item
  Montana State University Library\\
\item
  Funded the publication of \emph{Strand et al.~2024}
\end{itemize}

President's Prize - 2nd Place Speaker

2023

\begin{itemize}
\tightlist
\item
  Entomology 2023, National Harbor, MD\\
\item
  Entomological Society of America
\end{itemize}

Student Scholar Travel Award (\$500)

2023

\begin{itemize}
\tightlist
\item
  College of Agriculture, Montana State University
\end{itemize}

1st Place Speaker

2023

\begin{itemize}
\tightlist
\item
  2023 LRES GSO Student Research Colloquium\\
\item
  LRES Graduate Student Organization, Montana State University
\end{itemize}

Montana Grains Foundation Scholarship (\$2,500)

2022

\begin{itemize}
\tightlist
\item
  Montana Grain Growers Association
\end{itemize}

Gradaute Professional Advancement Grant (\$500)

2022

\begin{itemize}
\tightlist
\item
  The Graduate School, Montana State University
\end{itemize}

\subsection{Seminars \& Presentations}\label{seminars-presentations}

{Check out my~Talks \& Presentations~page}

Contributed presentations at scientific meetings

SE Sing, Strand JR, Peterson RKD, Weaver DK. 2025. Understanding Host
Preference in Mecinus janthinus and M. janthiniformis: Implications of
Hybridization for Effective Biological Control. National Forum on
Biological Control Conference. Annapolis, MD.

Strand JR, Sterling TM, Peterson RKD, Weaver DK. 2025. Agroecological
Importance of Smooth Brome in Hosting Braconid Parasitoids Associated
with the Insect Pest Wheat Stem Sawfly. Annual Meeting of the Western
Society of Weed Science. Seattle, WA. (poster)

Strand JR, Peterson RKD, Sterling TM, Weaver DK. 2024. Agroecological
importance of smooth brome in managing wheat stem sawfly (Hymenoptera:
Cephidae) via its associated braconid parasitoids. Session: Grad
Competition, Plant-Insect Ecosystems: IPM -- General and Field Crops.
Annual Meeting of the Entomological Society of America, Phoenix, AZ.

Strand JR, Weaver DK, Sterling TM, Peterson RKD. 2024. The role of
chemical ecology and agroecological importance of smooth brome in
biological control of the wheat stem sawfly by native parasitoids.
Masters of Entomology Defense. Montana State University, Bozeman, MT.

Sing SE, Ward SM, Peterson RKD, Strand JR, Friesenhahn E, Holland M,
Hubbard CE, Randall CB, Weaver DK. 2024. Biological control of hybrid
toadflax: insights from chemical ecology. National Forum on Biological
Control Conference. Annapolis, MD.

Strand JR. 2024. How smooth brome and parasitoids can aid in the
biological control of wheat stem sawfly. Montana Wheat and Barley
Committee - March Madness. Montana State University, Bozeman, MT.

Strand JR, Weaver DK, Sterling TM, Peterson RKD. 2023. Using organic
volatile compounds to assess the relationship between wheat stem sawfly,
parasitoids, and smooth brome. Session: Grad Competition, Plant-Insect
Ecosystems: Chemical Ecology 2. Annual Meeting of the Entomological
Society of America, National Harbor, MD. \textbf{Second place in
section}.

Strand JR, Weaver DK, Sterling TM, Peterson RKD. 2023. The importance of
smooth brome and parasitoids in wheat stem sawfly biocontrol. 2023 Land
Resources and Environmental Sciences Graduate Student Organization
Research Colloquium. Montana State University, Bozeman, MT.
\textbf{First place overall}.

Strand JR, Weaver DK, Sterling TM, Peterson RKD. 2022. Using organic
volatile compounds to assess the relationship between wheat stem sawfly,
parasitoids, and smooth brome. Session: Grad Competition, Plant-Insect
Ecosystems: Chemical Ecology 1. Annual Meeting of the Entomological
Society of America. Vancouver, BC, Canada.

Strand JR, Weaver DK, Sterling TM, Peterson RKD. 2022. Importance of
parasitoids and smooth brome for integrated pest management of wheat
stem sawfly. 2022 Land Resources and Environmental Sciences Graduate
Student Organization Research Colloquium. Montana State University,
Bozeman, MT.

Strand JR, Weaver DK, Sterling TM, Peterson RKD. 2021. Importance of
parasitoids and smooth brome for integrated pest management of wheat
stem sawfly. Session: Plant Insect-Ecosystems: IPM -- General. Annual
Meeting of the Entomological Society of America, Denver, CO.

Invited Guest Lectures

Strand JR. Can Bugs Help Us? ENSC 210: Role of Plants in the
Environment. Montana State University. Instructor: Will Wetzel. March
2025.

Strand JR. Introduction to Insect Identification and Sampling. BIOB 260:
Principals of Biological Diversity. Montana State University.
Instructors: Erin Teichroew and Chris Larson. August 2025.

Strand JR. Integrated Pest Management. ENSC 110: Land Resources and
Environmental Sciences. Montana State University. Instructor: Frank
Dougher. October 2025.

Workshops

Understanding Wheat Stem Sawfly. 2022 Integrated Pest Management
Workshop, ``MSU Extension IPM Workshop: Diagnosing Plant Health
Problems'' Montana State University, Bozeman, MT.

\subsection{Teaching}\label{teaching}

Teaching Assisstant

\textbf{Integrated Pest Management} (AGSC 401)\\
Fall 2023, 3 credits, 20 undergraduates, 10 graduate students\\
Online course taught by Dr.~David Weaver\\
Montana State Unviersity

\textbf{Integrated Pest Management} (AGSC 401)\\
Fall 2025, 3 credits, 25 undergraduates, 10 graduate students\\
Online course taught by Dr.~David Weaver\\
Montana State Unviersity

Invited Guest Lecturer

\textbf{Role of Plants in the Environment} (ENSC 210)\\
Spring 2025, 3 creidts\\
``Can Bugs Help Us?''\\
Course taught by Dr.~Will Wetzel\\
Montana State Unviersity

\textbf{Principals of Biological Diversity} (BIOB 260)\\
Fall 2025, 5 credits\\
``Introduction to Insect Identification and Sampling''\\
Course taught by Erin Teichroew and Chris Larson\\
Montana State Unviersity

\textbf{Land Resources and Environmental Sciences} (ENSC 110)\\
Fall 2025, 3 credits\\
``Integrated Pest Management''\\
Course taught by Frank Dougher\\
Montana State University

\subsection{Service}\label{service}

Peer Review Journal Activity

Journal of Environmental Entomology

2024 - present

Journal of Environmental Toxicology and Chemistry

2025 - present

Society Memberships

Montana State University Graduate Employee Organization (LRES Steward)

2025 - present

LRES Graduate Student Organization (Co-chair)

2022 - present

LRES Graduate Student Organization (Treasurer)

2021 - 2022

Entomological Society of America

2021 - present

Montana Grain Growers Association

2022 - present

Active Community Volunteer

Southwest Montana Mountain Bike Association

2021 - present

Gallatin Watershed Council

2021 - present

Gallatin Valley Land Trust

2021 - present

\subsection{Co-Hosted Events}\label{co-hosted-events}

LRES Graduate Student Organization Annual Symposium

LEPS 2025 Role: Chair\\
Location: Montana State University, Bozeman, MT\\
Date: April 7, 2025\\
Event website: LEPS 2025

2024 LRES-PSPP Student Research Symposium Role: Chair \& event emcee
Location: Montana State University, Bozeman, MT Date: April 18, 2024
Hyperlink: 2024 Symposium

2023 LRES Colloquium Role: Treasurer Location: Montana State University,
Bozeman, MT Date: April 13, 2023

2022 LRES Colloquium Role: Treasurer Location: Montana State University,
Bozeman, MT Date: April 29, 2022

\subsection{Employment History}\label{employment-history}

Pace Analytical - Onsite at 3M - Regulatory IT Analyst

2017 - 2021

(Maplewood, MN)

Provided digital support between 3M employees and the 3M Chemical Data
Management System (CDMS).

Integrated, tested, and supported user and data interface between CDMS
and SAP.

Generated, designed, and communicated unique, detailed reports of
material information using SQL, Microsoft Excel, and SAP.

Metropolitan Mosquito Control District - Seasonal Mosquito Technician

2017

(Oakdale, MN)

Worked on a team to collect and analyze mosquito larvae samples from
wetland areas across the St.~Paul metropolitan area.

Independently maintained low adult mosquito populations in Oakdale, MN
by surveying and treating both adults and larvae.

Leatherdale Equin Center - Barn Maintenance Crew

2015 - 2017

(St.~Paul, MN)\\

Maintained healthy conditions for equine patients at the University of
Minnesota Horse Veterinary Center and Large Animal Hospital.

Trained and managed new members of the crew.

Minnesota Youth Soccer Association - Youth Soccer Referee

2010 - 2021




\end{document}
